Type-Safe Modular Parsing

### 2016-12-13 paper/sections/Overview.tex
Progress

\huang{a paragraph here, refer to related work section, discuss
  previous works of modular parsing using parser generators and its
  drawbacks}\bruno{Please finish this.}

@@

Unfortunatelly many parser combinators have important limitations for
modular parsing. In particular several parser combinators,
including the famous library Parsec~\cite{} library, require
programmers to manually do\textit{left-recursion elimination}, and
require significant amounts \textit{backtracking}. Both of these are
problematic in a modular setting.

@@

\huang{todo: polish it}
\begin{lstlisting}
type Open[T] = (=> T) => T

@@

We take \lstinline{pApp} as an example again. It uses \lstinline{self} instead of hard-coded \lstinline{pExpr} for recursive parsing. Depending on the real argument, \lstinline{self} can be a parser which covers variables and applications, or a parser which also supports the lambda case.

\paragraph{Client Code} Since the type of parsing functions are changed, we can not apply the parsers directly. The \lstinline{fix} function below, which is the the standard \textit{fixpoint combinator}, is used to `close' the open recursion.

@@

\paragraph{Extensibility} As same as what we did in previous section, we extend the grammar by adding lambda abstractions as a new case. Notice that the old code is not touched, hence separate compilation is obtained.

@@

\huang{todo: rewrite}
In the new parser \lstinline{LambdaParser} extended from the old one,
we write a new parsing function \lstinline{pLam} for the new case
lambda abstraction. And compose it with \lstinline{pVarApp}, which we
already have in the old parser \lstinline{ExprParser}. Open recursion
is critical to achieve such type-safe compositions, because wherever
we refer to the parser of the whole expression, it is updated to cover
all of the three case we have now.

@@

Parsing with open recursion will be discussed further in section \ref{sec:openandparsing}.

>>>>

Modular parsing based on parser generators is supported by many libraries []. Users can separate the concrete/abstract syntax and related parsing code into reusable components. Then the corresponding parsers are built by their library utilities. For example, NOA [] uses Java annotation processing to collect grammar information, and then generates ANTLR4 parsers. However, such generation procedure requires a whole compilation after the collection of all grammar pieces. Once the grammar changed, even slightly, grammar information must be re-collected and the parser must be re-generated. Hence thoes libraries only have syntactic modularity.

Generating parsers often requires full information of grammars, thus semantic modularity is difficult to achieve in this way.

\huang{a paragraph here, refer to related work section, discuss
  previous works of modular parsing using parser generators and its
  drawbacks}\bruno{Please finish this.}\huang{done}

@@

Unfortunatelly many parser combinators have important limitations for
modular parsing. In particular several parser combinators,
including the famous library Parsec~\cite{} library, require
programmers to manually do \textit{left-recursion elimination}, and
require significant amounts \textit{backtracking}. Both of these are
problematic in a modular setting.

@@

\begin{lstlisting}
type Open[T] = (=> T) => T

@@

We take \lstinline{pApp} as an example again. It uses \lstinline{self} instead of hard-coded \lstinline{pExpr} for recursive parsing. Depending on the real argument, \lstinline{self} can be a parser which covers only variables and applications, or a parser which also supports the lambda case.

\paragraph{Client Code} Since the type of parsing functions are changed, we can not apply them directly. The client code below shows how to use the new parser. The \lstinline{fix} function, which is the the standard \textit{fixpoint combinator}, is used to `close' the open recursion.

@@

\paragraph{Extensibility} As same as what we did in previous section, we extend the grammar by adding lambda abstractions as a new case. Similarly, the new parser extends from the old one by adding \lstinline{pLam} on it. Notice that the old code is not touched, hence separate compilation is obtained.

@@

All recursive calls of parsing `an expression' are called via the explicit parameter \lstinline{self}. Once the \lstinline{fix} function is applied, these calls will be updated appropriately. Parsing functions such as \lstinline{pApp} do not need to be rewritten. As a consequence, this parser recognizes all valid expressions of the new grammar, including \lstinline{(\\x.x)} \lstinline{(\\y.y)} which cannot be parsed in the previous section.

@@

As we demonstrated, delegation encoded by open recursion is the key technique to obtain semantic modularity. It enables type-safe code reuse and separate compilation in our parsers. We will discuss this topic further in section \ref{sec:openandparsing}.
